---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment  = "#>",
                      dpi      = 300,
                      fig.path = "readme_resources/README-")
```

spec Package
============

This test package is meant to help reading and working with `.sig` formatted spectral
files in R.

It may become part of the spectral analysis workflow in the [Cavdender-Bares Lab]("https://www.cbs.umn.edu/research/research-cbs/faculty-labs/cavender").

---

As of now, you should probably install `spec` from the bitbucket repository. To
do, so use:

```{r, eval = FALSE}
library(devtools)

devtools::install_bitbucket(repo      = "cavenderbareslab/spec",
                            auth_user = "your_usr_name",
                            password  = "your_bitbucket_pwd"
                            )
```


Now you should be able to load the package simply by doing

```{r}
library(spec)
```


To read in sig files, you need to pass the full path of each file in a vector
to the function `read_spectra_sig`
```{r}
files   = dir("data/", full.names = T)

# File paths
files

# Read files! This can take some time...
spectra = spec::read_spectra_sig(files)
```


We may have taken a measurement on a white reference (pretending it was a leaf), just
to check if the real reference was still accurate (you should see a straight line).

Anyways, we can let `spec` know that we did so by annotating the `spectrum` with
a type `leaf` or `blank`. Type `leaf` is the default and the argument is reclycled. 

```{r}

# If all samples in files are the same type, you don't have to repeat yourself
spectra = spec::read_spectra_sig(files, types = "leaf")

# If samples are different types, you should pass a vector of types of the same
# length as files. Normally you'd be reading this info from a file.
sample_type = c("leaf", "blank")

# Hopefully this is TRUE
length(files) == length(sample_type)

# Now read the sig files
spectra = spec::read_spectra_sig(files, types = sample_type)
```


You are able to check out basic information / plot each individual spectrum or 
all the spectra. 

```{r}
# Check it out
spectra[[1]]

# Plot the first spectrum
plot(spectra[[1]])

# ...or all of them
plot(spectra)
```


I feel like the functions above are a quick and dirty way to check out if the 
`sig` files were read correctly. More often than not, you will want to work with
the spectral data in tabular format.

```{r}
spec_df = spec::convert_to_data_tables(spectra)
```


The `convert_to_data_tables` function returns a `list` with three `data.frames`:
    * `target`. With metadata and spectral values of the target measurement.
    * `reference`. Metadata and spectral values of the __unique__ reference measurements.
    * `reflectance`. The ratio between the `target` and `reference` measurements.

The `target$reference_id`

```{r}
spec_df$target[1:2, 1:30]

```



Known issues and TODOs
----------------------

* conversion to data frames will break if spectra have different lengths




